---
title: N170 from ERP-fix
author: Hannah, 9/12/2017
output:
  html_document:
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

```{r setup, include=FALSE, warnings=F}
knitr::opts_chunk$set(echo = FALSE)
options(width=140)
require(dplyr)
require(ggplot2)
require(lme4)
require(lmerTest)
require(tidyr)
```

**Total sample is 65 subjects**
*Changed to cross-classified models (not electrode nested within subject)*

### 1. N170 grand averages  
**Negative is plotted downwards.**

```{r plotGrand, echo=FALSE}
grand = read.delim("For_plotting_Cat_N170_noBS.txt")
grand$avgElec = (grand$TP7 + grand$TP8 + grand$P7 + grand$P8)/4

avg = select(grand, Time, Race, avgElec) %>% 
  group_by(Time, Race) %>% 
  summarise_each(funs(mean(., na.rm=TRUE))) %>% 
  as.data.frame()

N170box =  annotate("rect",    #N170
                    xmin=135, xmax=195, ymin=-Inf, ymax=Inf,
                    alpha=0,
                    fill="#F0E0FF",
                    color="black", 
                    linetype="dashed")

ERPline = geom_line(lwd=1.1, alpha=.85,
                    aes(color = Race, linetype = Race))

condLineType <- c("Black" = "solid",
                  "White" = "solid")

condColors <- c("Black" = "black", 
                "White" = "darkgreen")

none = element_blank() 

# average of all 9 electrodes
ggplot(data=avg, aes(Time, avgElec, group = Race)) + 
  ERPline + 
  N170box + 
  theme_bw() + 
  theme(panel.grid.major.x = none, panel.grid.minor.x = none) +
  scale_x_continuous("Time (ms)", 
                     limits=c(-100, 1000), 
                     expand=c(0,0),   # expand=c(0,0) removes extra space before & after data
                     breaks=c(-100, 0, 100, 200, 300, 400, 500, 600, 700, 800, 900, 1000)) +
  geom_hline(yintercept=0) + # adds x axis
  geom_vline(xintercept=0) +
#  scale_y_continuous(limits =c(10, -7.5)) +  # scale_y_reverse flips y axis
  ylab("Amplitude (uV)") +
  scale_color_manual(values=condColors) +
  scale_linetype_manual(values=condLineType) +
  ggtitle("Average of TP7, TP8, P7, P8") +
  theme(plot.title = element_text(hjust = 0.5)) # center title

# make different data frame to facet electrodes and plot them separately
forfacetting = select(grand, Time, Race, TP7, TP8, P7, P8) %>% 
  group_by(Time, Race) %>% 
  summarise_each(funs(mean(., na.rm=TRUE))) %>% 
  as.data.frame() %>% 
  gather("Electrode", "MeanAmp", 3:6)

# each electrode separately
ggplot(data=forfacetting, aes(Time, MeanAmp, group = Race)) + 
  ERPline + 
  N170box + 
  theme_bw() + 
  facet_wrap(~Electrode) +
  theme(panel.grid.major.x = none, panel.grid.minor.x = none) +
  scale_x_continuous("Time (ms)", 
                     limits=c(-100, 600), 
                     expand=c(0,0),   # expand=c(0,0) removes extra space before & after data
                     breaks=c(-100, 0, 100, 200, 300, 400, 500, 600)) +
  geom_hline(yintercept=0) + # adds x axis
  geom_vline(xintercept=0) +
#  scale_y_continuous(limits =c(10, -7.5)) +  # scale_y_reverse flips y axis
  ylab("Amplitude (uV)") +
  scale_color_manual(values=condColors) +
  scale_linetype_manual(values=condLineType) +
#  ggtitle("Average of TP7, TP8, P7, P8") +
  theme(plot.title = element_text(hjust = 0.5),# center title
        axis.title = element_text(size = 20),
        axis.text.y = element_text(size = 16),
        axis.text.x = element_text(size = 12),
        legend.title = element_text(size = 20),
        legend.text = element_text(size = 16),
        strip.text = element_text(size = 16)) 

ggsave("./FacettedGrandAverages_N170.tiff")

```
  
####Signal averaging approach:

All trials were averaged together and the mean amplitude was calculated for each condition for each subject. Subject and electrode were used as crossed random factors. ICCs were first calculated with an intercept-only model. 

``` {r}
dat = read.delim("./Cat_AllSubs_TBTaverages_noBS_groupN170.txt")

sigavg = select(dat, -Trial, -Condition, -Fix) %>% 
  group_by(Subject, Race, Electrode) %>% 
  summarise_each(funs(mean(., na.rm = T))) %>% 
  as.data.frame()

sum = lmer(MeanAmp ~ (1|Subject) + (1|Electrode), dat = sigavg) %>% summary()

subICC = 4.8994/(4.8994+.3832+3.0517)
elecICC = .3832/(4.8994+.3832+3.0517)
```

The ICC associated with *subject* is `r subICC`.  
The ICC associated with *electrode* is `r elecICC`.

A model was fitted with Race predicting mean amplitude calculated in this way. The intercept and slope of Race were allowed to vary by subject. The intercept was allowed to vary by Electrode. 

Race was effect coded. 

``` {r test, echo=F}
# add effect codes for categorical variables
sigavg$Race.e = NA
sigavg$Race.e[sigavg$Race == "Black"] = -1
sigavg$Race.e[sigavg$Race == "White"] = 1

grand1 = lmer(MeanAmp ~ Race.e + (Race.e|Subject) + (1|Electrode), dat = sigavg)
```
  
**Random effects:**  
``` {r grand1random, echo = FALSE}
summary(grand1)$varcor
```

**Fixed effects:**
``` {r grand1fixed, echo = FALSE} 
round(summary(grand1)$coefficients, digits = 3)
```

### 2. Looking at the N170 over the course of the experiment

ICCs were first calculated for subject and electrode.
``` {r}
sum = lmer(MeanAmp ~ (1|Subject) + (1|Electrode), dat = dat) %>% summary()

subICC = 5.1535/(5.1535+.4853+101.8823)
elecICC = .4853/(5.1535+.4853+101.8823)
```

The ICC associated with *subject* is `r subICC`.  
The ICC associated with *electrode* is `r elecICC`.  


**Slopes and estimates of lines are from the MLM, not fitted with OLS.  Negative is plotted downward.** 

``` {r trialLevel, echo = FALSE, warning = FALSE}
# making dummy codes is redundant but comforting, so LET ME LIVE
dat$Race.d = NA
dat$Race.d[dat$Race == "Black"] = 0
dat$Race.d[dat$Race == "White"] = 1

# find slopes and intercepts with dummy coded model, trial (unscaled) is included
plot1.d = lmer(MeanAmp ~ Race.d*Trial + (Race.d|Subject) + (1|Electrode), dat = dat)

parms <- fixef(plot1.d) # fixed parameters from model
vcmat <- vcov(plot1.d) # variance/covariance matrix of fixed effects

# First calculate slopes

## each row of this matrix defines entries of parms that we want to combine:
STmat <- matrix(NA,2,4)
STmat[1,] <- c(0,0,1,0) # slope for race = 0 (black)
STmat[2,] <- c(0,0,1,1) # slope for race = 1 (white)

Sparest <- STmat %*% parms          # see above for notation

# Next calculate intercepts

## each row of this matrix defines entries of parms that we want to combine:
ITmat <- matrix(NA,2,4)
ITmat[1,] <- c(1,0,0,0) # intercept for race = 0 (black)
ITmat[2,] <- c(1,1,0,0) # intercept for race = 1 (white)

Iparest <- ITmat %*% parms          # see above for notation

forPlotting = cbind(Sparest, Iparest) %>% as.data.frame()
names(forPlotting) = c("Slope", "Intercept")

# label for understanding
forPlotting$Race = c("Black", "White")

# calculate SE for each trial, separately for each race
SETmat <- matrix(NA,512,4)
SETmat[,1] = 1 # first column is intercept
SETmat[1:256, 2] = 0 # second column is race. For first half of rows, race = 0 (Black) 
SETmat[1:256, 3] = 1:256 # third column is trial. Different row for each trial
SETmat[1:256, 4] = 0 # fourth column is interaction. 0 when race = 0, trial number when race = 1
SETmat[257:512, 2] = 1 # second column is race. For second half of rows, race = 1 (White) 
SETmat[257:512, 3] = 1:256 # third column is trial. Different row for each trial
SETmat[257:512, 4] = 1:256 # fourth column is interaction. 0 when race = 0, trial number when race = 1

newvc <- SETmat %*% vcmat %*% t(SETmat)
ses <- sqrt(diag(newvc)) # long string of SEs for each trial, first 256 are for Black faces, second 256 are for White faces

SE.dat = data.frame(Race = rep(c("Black", "White"), each=256), Trial = rep(c(1:256), 2), SE = ses)

# sloppy, and there's probably a better way of doing this
# but take slope and intercept and calculate all the points on the line for each trial so that I can calculate upper and lower 95% CIs 
SE.dat$Point = NULL
SE.dat$Point[SE.dat$Race == "Black"] = forPlotting$Intercept[forPlotting$Race == "Black"] + 1:256*forPlotting$Slope[forPlotting$Race == "Black"]
SE.dat$Point[SE.dat$Race == "White"] = forPlotting$Intercept[forPlotting$Race == "White"] + 1:256*forPlotting$Slope[forPlotting$Race == "White"]

# calculate upper and lower bound (+/- 1 SE)
SE.dat$upper = SE.dat$Point + SE.dat$SE
SE.dat$lower = SE.dat$Point - SE.dat$SE

# BINGO -------------------------------------------------------------------
ggplot(SE.dat, aes(Trial, Point, alpha = Race, color = Race, shape = Race)) +
  geom_abline(data = forPlotting, aes(intercept=Intercept, slope=Slope, color = Race), size=1)+
  geom_ribbon(aes(ymin=lower, ymax=upper, x = Trial),
              linetype = "dashed",
              alpha = .1,
              color = "black") +
  labs(x = "Trial", y = "N170 Mean Amplitude") +
  #  scale_shape_manual(values=c(1,19,1,19)) +
  scale_alpha_manual(values=c(.1,.1,.1,.1)) +
  scale_color_manual(values=c("black", "darkgreen")) +
  theme_bw() +
  scale_y_continuous(limits =c(-3, .5), expand=c(0,0)) +
  scale_x_continuous(expand=c(0,0)) +
  #  ggtitle("Trial is original trial number") +
  theme(plot.title = element_text(hjust = 0.5),# center title
        axis.title = element_text(size = 20),
        axis.text = element_text(size = 16),
        legend.title = element_text(size = 20),
        legend.text = element_text(size = 16)) 

ggsave("./SimpleSlopes_N170.tiff")
```

##### Simple slopes

Trial is scaled to range from 0 to 10 (instead of 1 to 256) so that the betas associated with trial are a little bigger (but significance testing is unaffected by linear scaling, so the test statistics and p values will be the same as if we used the unscaled Trial variable).

``` {r simple, echo = FALSE}
# rescale trial
dat$Trial.begin = (dat$Trial-1)/25.5
# shift trial to look at fixed effects at end of task as well
dat$Trial.end = dat$Trial.begin - 10

# find slopes with dummy coded model, trial is scaled
begin.d <- lmer(MeanAmp ~ Race.d*Trial.begin + (Race.d|Subject) + (1|Electrode), dat = dat)

parms <- fixef(begin.d) # fixed parameters from model
vcmat <- vcov(begin.d) # variance/covariance matrix of fixed effects

## each row of this matrix defines entries of parms that we want to combine:
Tmat <- matrix(NA,2,4)
Tmat[1,] <- c(0,0,1,0) # slope for race = 0 (black)
Tmat[2,] <- c(0,0,1,1) # slope for race = 1 (white)

parest <- Tmat %*% parms          # see above for notation
newvc <- Tmat %*% vcmat %*% t(Tmat)
ses <- sqrt(diag(newvc)) 

## final results
fin = cbind(parest, ses) %>% as.data.frame()
names(fin) = c("est", "ses")

# to calculate 95% CI intervals, lower bound = m - 2*SE, upper bound = m + 2*SE

fin$lbnd = fin$est - 2*fin$ses
fin$ubnd = fin$est + 2*fin$ses

fin = format(fin, digits = 3)

# relabel for understanding
fin$Race = c("Black", "White")

fin = rename(fin, Estimate = est, SE = ses, ci95_lower = lbnd, ci95_upper = ubnd)

# display
fin
```  

#### Model output: Beginning of experiment
The intercept, slopes of current and previous trial condition and their interaction are allowed to vary by subject. *Categorical variables are effect coded.*   
  
Trial is scaled to range from 0 to 10.  

``` {r begin, echo = FALSE}
# same model, but with effect coding
dat$Race.e = NA
dat$Race.e[dat$Race == "Black"] = -1
dat$Race.e[dat$Race == "White"] = 1

begin.e = lmer(MeanAmp ~ Race.e*Trial.begin + (Race.e|Subject) + (1|Electrode), dat = dat)
```
  
**Random effects:**  
``` {r beginrandom, echo = FALSE}
summary(begin.e)$varcor
```

**Fixed effects:**
``` {r beginfixed, echo = FALSE} 
round(summary(begin.e)$coefficients, digits = 3)
```

#### Model output: End of experiment
Trial is scaled to range from 10 to 0.  

``` {r end, echo = FALSE}
# same model, but with effect coding
end.e = lmer(MeanAmp ~ Race.e*Trial.end + (Race.e|Subject) + (1|Electrode), dat = dat)
```
  
**Random effects:**  
``` {r endrandom, echo = FALSE}
summary(end.e)$varcor
```

**Fixed effects:**
``` {r endfixed, echo = FALSE} 
round(summary(end.e)$coefficients, digits = 3)
```

### 3. Testing variances of slopes and intercepts using log likelihood tests
``` {r}
# testing significances of random intercepts/slopes
m3 = lmer(MeanAmp ~ Race.d*Trial.begin + (1|Electrode), dat = dat)
m2 = lmer(MeanAmp ~ Race.d*Trial.begin + (1|Subject), dat = dat)
m1 = lmer(MeanAmp ~ Race.d*Trial.begin + (Race.d|Subject), dat = dat)
m0 = lm(MeanAmp ~ Race.d*Trial.begin, dat = dat)

# test intercept of subject
anova(m2, m0)

# test slope of race by subject
anova(m2, m1)

# test intercept of electrode
anova(m3, m0)

```
