---
title: N170 from ERP-fix
author: Hannah, 4/4/2017
output:
  html_document:
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

```{r setup, include=FALSE, warnings=F}
knitr::opts_chunk$set(echo = TRUE)
options(width=140)
require(dplyr)
require(ggplot2)
require(lme4)
require(lmerTest)
require(colorspace)
require(grid)
require(tidyr)
```

**Total sample is 134 subjects, 60 from CU and 74 from MU.**

### 1. N170 grand averages  
**Negative is plotted downwards.**

```{r plotGrand, echo=FALSE}
grand = read.delim("For_plotting_Cat_N170_noBS.txt")
grand$avgElec = (grand$TP7 + grand$TP8 + grand$P7 + grand$P8)/4

avg = select(grand, Time, Race, avgElec) %>% 
  group_by(Time, Race) %>% 
  summarise_each(funs(mean(., na.rm=TRUE))) %>% 
  as.data.frame()

N170box =  annotate("rect",    #N170
                    xmin=135, xmax=195, ymin=-Inf, ymax=Inf,
                    alpha=0,
                    fill="#F0E0FF",
                    color="black", 
                    linetype="dashed")

ERPline = geom_line(lwd=1.1,
                    linetype="solid",
                    aes(color = Race))

condColors <- c("Black" = "red3", 
                "White" = "dodgerblue4")

none = element_blank() 

ERPline = geom_line(lwd=1.1,
                    aes(color = Race))



none = element_blank() 

# average of all 9 electrodes
ggplot(data=avg, aes(Time, avgElec, group = Race)) + 
  ERPline + 
  N170box + 
  theme_bw() + 
  theme(panel.grid.major.x = none, panel.grid.minor.x = none) +
  scale_x_continuous("Time (ms)", 
                     limits=c(-100, 1000), 
                     expand=c(0,0),   # expand=c(0,0) removes extra space before & after data
                     breaks=c(-100, 0, 100, 200, 300, 400, 500, 600, 700, 800, 900, 1000)) +
  geom_hline(yintercept=0) + # adds x axis
  geom_vline(xintercept=0) +
#  scale_y_continuous(limits =c(10, -7.5)) +  # scale_y_reverse flips y axis
  ylab("Amplitude (uV)") +
  scale_color_manual(values=condColors) +
  ggtitle("Average of TP7, TP8, P7, P8") +
  theme(plot.title = element_text(hjust = 0.5)) # center title

# make different data frame to facet electrodes and plot them separately
forfacetting = select(grand, Time, Race, TP7, TP8, P7, P8) %>% 
  group_by(Time, Race) %>% 
  summarise_each(funs(mean(., na.rm=TRUE))) %>% 
  as.data.frame() %>% 
  gather("Electrode", "MeanAmp", 3:6)

# each electrode separately
ggplot(data=forfacetting, aes(Time, MeanAmp, group = Race)) + 
  ERPline + 
  N170box + 
  theme_bw() + 
  facet_wrap(~Electrode) +
  theme(panel.grid.major.x = none, panel.grid.minor.x = none) +
  scale_x_continuous("Time (ms)", 
                     limits=c(-100, 600), 
                     expand=c(0,0),   # expand=c(0,0) removes extra space before & after data
                     breaks=c(-100, 0, 100, 200, 300, 400, 500, 600)) +
  geom_hline(yintercept=0) + # adds x axis
  geom_vline(xintercept=0) +
#  scale_y_continuous(limits =c(10, -7.5)) +  # scale_y_reverse flips y axis
  ylab("Amplitude (uV)") +
  scale_color_manual(values=condColors) +
#  ggtitle("Average of TP7, TP8, P7, P8") +
  theme(plot.title = element_text(hjust = 0.5)) # center title


```
  
To test the mean amplitude of the N170, a model was fitted with Race as a predictor. The intercept and slope of Race were allowed to vary by subject. The intercept was allowed to vary by Electrode nested within Subject. 

Race was effect coded. 

``` {r test, echo=F}
dat = read.delim("./Cat_AllSubs_TBTaverages_noBS_groupN170.txt")

# add effect codes for categorical variables
dat$Race.e = NA
dat$Race.e[dat$Race == "Black"] = -1
dat$Race.e[dat$Race == "White"] = 1

# model specified in "6 ERN model specification.R"
grand1 = lmer(MeanAmp ~ Race.e + (Race.e|Subject) + (1|Electrode:Subject), dat = dat)
```
  
**Random effects:**  
``` {r grand1random, echo = FALSE}
summary(grand1)$varcor
```

**Fixed effects:**
``` {r grand1fixed, echo = FALSE} 
round(summary(grand1)$coefficients, digits = 3)
```

### 2. Looking at the N170 over the course of the experiment

**Slopes and estimates of lines are from the MLM, not fitted with OLS.  Negative is plotted downward.** 

``` {r trialLevel, echo = FALSE, warning = FALSE}
# making dummy codes is redundant but comforting, so LET ME LIVE
dat$Race.d = NA
dat$Race.d[dat$Race == "Black"] = 0
dat$Race.d[dat$Race == "White"] = 1

# find slopes and intercepts with dummy coded model, trial (unscaled) is included
plot1.d = lmer(MeanAmp ~ Race.d*Trial + (Race.d|Subject) + (1|Electrode:Subject), dat = dat)

parms <- fixef(plot1.d) # fixed parameters from model
vcmat <- vcov(plot1.d) # variance/covariance matrix of fixed effects

# First calculate slopes

## each row of this matrix defines entries of parms that we want to combine:
STmat <- matrix(NA,2,4)
STmat[1,] <- c(0,0,1,0) # slope for race = 0 (black)
STmat[2,] <- c(0,0,1,1) # slope for race = 1 (white)

Sparest <- STmat %*% parms          # see above for notation
Snewvc <- STmat %*% vcmat %*% t(STmat)
Sses <- sqrt(diag(Snewvc)) 

slopes = cbind(Sparest, Sses) %>% as.data.frame()
names(slopes) = c("Slope", "Slope_SE")

# Next calculate intercepts

## each row of this matrix defines entries of parms that we want to combine:
ITmat <- matrix(NA,2,4)
ITmat[1,] <- c(1,0,0,0) # intercept for race = 0 (black)
ITmat[2,] <- c(1,1,0,0) # intercept for race = 1 (white)

Iparest <- ITmat %*% parms          # see above for notation
Inewvc <- ITmat %*% vcmat %*% t(ITmat)
Ises <- sqrt(diag(Inewvc)) 

intercepts = cbind(Iparest, Ises) %>% as.data.frame()
names(intercepts) = c("Intercept", "Intercept_SE")

forPlotting = cbind(slopes, intercepts)

# label for understanding
forPlotting$Race = c("Black", "White")


# BINGO -------------------------------------------------------------------
ggplot(dat, aes(Trial, MeanAmp, alpha = Race, color = Race, shape = Race)) +
  geom_point() +
  geom_abline(data = forPlotting, aes(intercept=Intercept, slope=Slope, color = Race), size=1)+
  labs(x = "Trial", y = "Mean Amplitude") +
#  scale_shape_manual(values=c(1,19,1,19)) +
  scale_alpha_manual(values=c(.1,.1,.1,.1)) +
  scale_color_manual(values=c("blue", "dodgerblue")) +
  theme_bw() +
  scale_y_continuous(limits =c(-5, 3), expand=c(0,0)) +
  scale_x_continuous(expand=c(0,0)) +
#  ggtitle("Trial is original trial number") +
  theme(plot.title = element_text(hjust = 0.5)) # center title
```

##### Simple slopes

Trial is scaled to range from 0 to 10 (instead of 1 to 256) so that the betas associated with trial are a little bigger (but significance testing is unaffected by linear scaling, so the test statistics and p values will be the same as if we used the unscaled Trial variable).

``` {r simple, echo = FALSE}
# rescale trial
dat$Trial.begin = (dat$Trial-1)/25.5
# shift trial to look at fixed effects at end of task as well
dat$Trial.end = dat$Trial.begin - 10

# find slopes with dummy coded model, trial is scaled
begin.d <- lmer(MeanAmp ~ Race.d*Trial.begin + (Race.d|Subject) + (1|Electrode:Subject), dat = dat)

parms <- fixef(begin.d) # fixed parameters from model
vcmat <- vcov(begin.d) # variance/covariance matrix of fixed effects

## each row of this matrix defines entries of parms that we want to combine:
Tmat <- matrix(NA,2,4)
Tmat[1,] <- c(0,0,1,0) # slope for race = 0 (black)
Tmat[2,] <- c(0,0,1,1) # slope for race = 1 (white)

parest <- Tmat %*% parms          # see above for notation
newvc <- Tmat %*% vcmat %*% t(Tmat)
ses <- sqrt(diag(newvc)) 

## final results
fin = cbind(parest, ses) %>% as.data.frame()
names(fin) = c("est", "ses")

# to calculate 95% CI intervals, lower bound = m - 2*SE, upper bound = m + 2*SE

fin$lbnd = fin$est - 2*fin$ses
fin$ubnd = fin$est + 2*fin$ses

fin = format(fin, digits = 3)

# relabel for understanding
fin$Race = c("Black", "White")

fin = rename(fin, Estimate = est, SE = ses, ci95_lower = lbnd, ci95_upper = ubnd)

# display
fin
```  

#### Model output: Beginning of experiment
The intercept, slopes of current and previous trial condition and their interaction are allowed to vary by subject. *Categorical variables are effect coded.*   
  
Trial is scaled to range from 0 to 10.  

``` {r begin, echo = FALSE}
# same model, but with effect coding
begin.e = lmer(MeanAmp ~ Race.e*Trial.begin + (Race.e|Subject) + (1|Electrode:Subject), dat = dat)
```
  
**Random effects:**  
``` {r beginrandom, echo = FALSE}
summary(begin.e)$varcor
```

**Fixed effects:**
``` {r beginfixed, echo = FALSE} 
round(summary(begin.e)$coefficients, digits = 3)
```

#### Model output: End of experiment
Trial is scaled to range from 10 to 0.  

``` {r end, echo = FALSE}
# same model, but with effect coding
end.e = lmer(MeanAmp ~ Race.e*Trial.end + (Race.e|Subject) + (1|Electrode:Subject), dat = dat)
```
  
**Random effects:**  
``` {r endrandom, echo = FALSE}
summary(end.e)$varcor
```

**Fixed effects:**
``` {r endfixed, echo = FALSE} 
round(summary(end.e)$coefficients, digits = 3)
```

